### 2018-08-29 submit/paper.tex

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\title{\huge Cross-language Sentiment Analysis for Social Media}

\author{\IEEEauthorblockN{Boyang Yan}
\IEEEauthorblockA{\textit{School of Computing and Information Technology} \\
  \textit{University of Wollongong}\\
  Wollongong, Australia\\
  by932@uowmail.edu.au}
\and
\IEEEauthorblockN{Xiaoxia Pu}
\IEEEauthorblockA{\textit{School of Computing and Information Technology} \\
  \textit{University of Wollongong}\\
  Wollongong, Australia\\
  xp816@uowmail.edu.au
}
\and
\IEEEauthorblockN{Xudong Zhang}
\IEEEauthorblockA{\textit{School of Computing and Information Technology} \\
  \textit{University of Wollongong}\\
  Wollongong, Australia\\
  xz944@uowmail.edu.au}
\and
\IEEEauthorblockN{Helene Tran}
\IEEEauthorblockA{\textit{School of Computing and Information Technology} \\
  \textit{University of Wollongong}\\
  Wollongong, Australia\\
  ht185@uowmail.edu.au}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
sentiment analysis, machine translation
\end{IEEEkeywords}

\section{Introduction}
Sentiment analysis is a part of text data mining. The aim of sentiment analysis
is to determine the attitude of speakers or writers with respect to some topic
or the overall contextual polarity or emotional reaction to a text document. It is usually equated with
opinion mining, which involves the use of natural language processing and
machine learning to ascertain the possibility of positive or negative opinions \cite{sentimentAnalysis}.
Sentiment analysis is useful for analyzing a huge amount of data relating to personal
opinions. It can be used in an e-business context. For example, business managers can analysise
customers' attitudes, as to whether they like or dislike their product or service.
Also, government can use sentiment analysis to analyze citizen perspectives.
In a word, sentiment analysis is coming into widespread use.

As Dr. Haiyun mentions, English language sentiment analysis research has
undergone major developments in recent years \cite{ChineseSentimentAnalysis}.
However, less research has been undertaken in other languages, such as Chinese.
Today, a lot of English language sentiment analysis theories have been
developed. Also, lots of Machine Translation tools are available, such as Google
translation, Bing translation and Yandex translation.
Manchine Translation uses computational linguistic programs and natural language
processing theory \cite{machineTranslation}.

However, nobody working in a combination of these two fields of research has undertaken non-English
sentiment analysis. Therefore, the research described and discussed in this
document aimed to create a testing model to find the best-combination of English
sentiment analysis tools and machine translation tools to obtain reliable
sentiment analysis results from non-English texts, for non-English speakers who
do not have such sentiment analysis tools to analyze their own language.
\section{Literature Review}
This research consists of three components.
Firstly, measuring machine translation service quality.
Secondly, testing sentiment analysis service quality.
Thirdly, finding the best compound mode for machine translation service and
sentiment analysis service.
As a result, this section will fouced on machine translation and sentiment
analysis.

\subsection{Machine Translation}
There are two research articles about testing modeling of machine translation.

\paragraph{Around Trip Translation Method}
\cite{roundTripTranslation}
As Somers argue that they establish an around trip translation method for detect
the quality of machine translation.
For example, testing English to Chinese translation tool.
Firstly, using English to Chinese translation tool translate test data to
Chinese.
Secondly, translation Chinese translated data back to English.
Finally, compare the similarity for two English data set.

\paragraph{Translation Method}
Another article is about using third-party language to test the quality of machine
translation (Pesu, 2017). For example, if testing English to Chinese translation
tool. Firstly, random choice an Intermediate third-party language. Secondly,
translation English test data to the third-party language, after translation
third-party language to Chinese. These two steps are only finished one path.
Another path is a translation in English directly to Chinese. In the end, the
comparison the two paths of results similarity. In this article, the main
finding is Google Translate is the best machine translation. In addition, there
is a tendency to be produced better results in European languages, which use
ANOVA Statistics method getting this conclusion. In my experiment, I also got
Google Translator is the best machine translation compare with Yandex and Baidu.

The three different machine translation testing modeling, we can call the three
different Metamorphic Relations. So, I have found a research article(Cao,2013)
about the effectiveness of metamorphic relations.

Also, when I finished reading all of the details, I have found Two-way ANOVA is also suitable for part 3 of my research (finding the best compound mode for machine translation service and sentiment analysis service).
I think that the disadvantages of those three-testing modeling will be involved some of the noise.
 It is clear the Round-Trip will be involved translation back path noise. Using the third language will be involved in third language translation path noise. My modeling will be involved sentiment analysis tools’ noise. However, my modeling does not involve any noise from translation tool. I think my modeling is better than the other two testing modeling. I am total from another aspect to test translation tool.

\section{Method}
\section{Test Data}
Total have 46180 movies reviews.
\begin{tabular}{lrl}
Ranking & Number of Test Data & Percentage\\
\hline
Ranking 10 & 7353 & 15.92 \%\\
Ramking 20 & 11209 & 24.27 \%\\
Ranking 30 & 16223 & 35.13 \%\\
Ranking 40 & 7663 & 16.59 \%\\
Ranking 50 & 3732 & 8.08 \%\\
\end{tabular}

\section{IQR, Mean, Median, Q1, Q3, lower extreme, upper extreme, mean slope, median slope}
\subsection{Baidu Chinese Sentiment analysis Positive Probability Base On (Chinese Origin Data)}
\begin{tabularx}{\textwidth}{ |X|X|X|X|X|X|X|X }
Ranking & IQR & Mean & Median & Q1 & Q3 & lowerExtreme & upperExtreme\\
\hline
10 & 0.3426533 & 0.2404781 & 0.1789910 & 0.0489987 & 0.3916520 & -0.4649812 & 0.9056319\\
20 & 0.3817485 & 0.2949725 & 0.2489580 & 0.0878135 & 0.4695620 & -0.4848092 & 1.0421847\\
30 & 0.4504405 & 0.3980354 & 0.3808120 & 0.1516095 & 0.6020500 & -0.5240512 & 1.2777107\\
40 & 0.5255830 & 0.5123376 & 0.5385910 & 0.2461685 & 0.7717515 & -0.5422060 & 1.5601260\\
50 & 0.5368080 & 0.5712128 & 0.6188395 & 0.3101650 & 0.8469730 & -0.4950470 & 1.6521850\\
\end{tabularx}

\begin{tabular}{rr}
  \hline
  Mean Slope & Median Slope \\
  \hline
  0.008788344 & 0.0116933 \\
  \hline
\end{tabular}

\bibliographystyle{IEEEtran}
\bibliography{library}
\end{document}